---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# almanac

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/DavisVaughan/almanac.svg?branch=master)](https://travis-ci.org/DavisVaughan/almanac)
[![Codecov test coverage](https://codecov.io/gh/DavisVaughan/almanac/branch/master/graph/badge.svg)](https://codecov.io/gh/DavisVaughan/almanac?branch=master)
<!-- badges: end -->

```{r}
library(almanac)
```

almanac implements a _grammar of schedules_, providing the fundamental building blocks to construct recurrence rules that identify "events" such as weekends or holidays. Constructing recurrence rules looks a little like this:

```{r}
# Thanksgiving = "The fourth Thursday in November"
on_thanksgiving <- yearly() %>% 
  recur_on_ymonth("November") %>%
  recur_on_wday("Thursday", nth = 4)
```

After constructing a recurrence rule, it can be used to generate dates that are in the "recurrence set".

```{r}
sch_seq("2000-01-01", "2006-01-01", on_thanksgiving)
```

Or determine if a particular date is a part of the recurrence set.

```{r}
sch_in(c("2000-01-01", "2000-11-23"), on_thanksgiving)
```

It also allows you to shift an existing sequence of dates, "stepping over" dates that are in the recurrence set.

```{r}
wednesday_before_thanksgiving <- "2000-11-22"

# Step forward 2 non-event days, stepping over thanksgiving
sch_step(wednesday_before_thanksgiving, n = 2, on_thanksgiving)
```

## Learning More

View the vignettes on [the website](https://davisvaughan.github.io/almanac/index.html) to learn more about how to use almanac.

- `vignette("almanac")`

- `vignette("adjust-and-shift")`

## Installation

You can NOT install the released version of almanac from [CRAN](https://CRAN.R-project.org) with:

``` r
# NO! install.packages("almanac")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DavisVaughan/almanac")
```

