---
title: "Introduction to almanac"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{almanac}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette is designed to introduce some of the common terminology used in {almanac} to get you up to speed on how to use the package. Along the way, we will see example usage of a number of the building blocks that will allow you to construct more complex schedules.

```{r setup}
library(almanac)
```

## Recurrence Rules

A _recurrence rule_ is a structured object that determines if a date should be counted as an _event_ or not. At the most basic level, the job of a recurrence rule is to return `TRUE` if a particular date is an _event_, and `FALSE` otherwise.

To build a recurrence rule, you start with a base recurrence frequency. There are 4 frequencies to choose from:

- `daily()`

- `weekly()`

- `monthly()`

- `yearly()`

Take the `yearly()` frequency, for example. By default, this will declare that 1 value per year is an event.

```{r}
on_yearly <- yearly()
on_yearly
```

The return value of `yearly()` is a _rrule_ object, short for "recurrence rule". This base object is all you need to start doing something useful. For example, you can provide `sch_seq()` this recurrence rule along with a `from` and `to` date, and it will return the events that it finds in the date range.

```{r}
sch_seq("1990-01-01", "1995-12-31", on_yearly)
```

What if we want a yearly value, but we want it on January 5th every year, rather than on the 1st? `yearly()` has a _very_ important argument called `since` that controls two things: the start date of the recurrence rule, and information such as the month, or the day of the month to use for the event dates if no other conditions have been specified to override those.

The default of `since` is set to `1970-01-01`, the Unix epoch, but this is arbitrary. It is because of this default that in the above example with `sch_seq()` we get values on January 1st. Let's change that.

```{r}
on_yearly_jan_5 <- yearly(since = "1990-01-05")

sch_seq("1990-01-01", "1995-12-31", on_yearly_jan_5)
```

__Notice that we also set the `since` date to be closer to our range of interest. This can dramatically improve performance in some cases.__

Now that the `since` date has been set to 1990, if we try and find yearly dates before 1990, they will not be included.

```{r}
# same as above, because the 1988 and 1989 dates are not included. 
sch_seq("1988-01-01", "1995-12-31", on_yearly_jan_5)
```

## Recurrence Set

Each recurrence rule comes with its own unique _recurrence set_. A recurrence set is the set of all event dates that are included in that rule. In the previous example, we used `sch_seq()` to extract a subset of dates from the recurrence set that were between `from` and `to`. You can also check if an existing date is included in a recurrence rule's set with `sch_in()`.

```{r}
on_monthly <- monthly(since = "1990-01-10")

x <- as.Date("2000-01-08") + 0:5
x

x_in_set <- sch_in(x, on_monthly)
x_in_set

x[x_in_set]
```

### Caching

{almanac} makes an attempt to be smart by caching dates in the recurrence set that it has "seen". This means that repeated access to the same schedule can often speed up substantially between the first access and subsequent accesses.

```{r}
start_date <- "1990-01-01"
on_daily <- daily(since = start_date)

# The first time is "slow"
system.time(sch_seq(start_date, "2000-01-01", on_daily))

# Repeated access to the same location is fast
system.time(sch_seq(start_date, "2000-01-01", on_daily))

# Dates are cached from `[since, to]`, so if you back up `to`, its still fast
system.time(sch_seq(start_date, "1900-05-01", on_daily))

# But when you extend `to` forward in time, it will have to generate all dates
# between `since` and `to` again
system.time(sch_seq(start_date, "2000-05-01", on_daily))
```

## Recurrence Conditions

So far we have worked with the base recurrence rules. Things get _much_ more interesting when we start adding extra conditions to these rules. Conditions are a way to _limit_ or _expand_ a given recurrence rule to hone in on recurring dates that you are particularly interested in. All condition functions in {almanac} start with `recur_*()`. For example, let's take a monthly rule, which defaults to give us 1 day per month, and _expand_ it to give us every 4th and 16th day of the month.

```{r}
on_4th_and_16th <- monthly(since = "2000-01-01") %>%
  recur_on_mday(c(4, 16))

sch_seq("2000-01-01", "2000-06-01", on_4th_and_16th)
```

An important thing to note here is that even though our `since` date is on the first of the month, we are "overriding" that with the recurrence condition, so that information is not used.

Recurrence rules can continually be added to further refine your rule. When you add a condition to a rule, you get another rule back. Let's try creating a rule for the recurring holiday, Labor Day. This recurs on the first Monday of September, yearly. To do this, we will:

- Use a `yearly()` base since this happens 1 time per year.

- Use `recur_on_ymonth()` to hone in on September.

- Use `recur_on_wday()` to hone in on the first Monday of the month.

```{r}
on_labor_day <- yearly(since = "2000-01-01") %>%
  recur_on_ymonth("Sep") %>%
  recur_on_wday("Monday", nth = 1)

sch_seq("2000-01-01", "2005-01-01", on_labor_day)
```

The `nth` argument of `recur_on_wday()` is especially useful for selecting from the end of the month. If we wanted the _last_ Monday in September instead, we could do:

```{r}
on_last_monday_in_sept <- yearly(since = "2000-01-01") %>%
  recur_on_ymonth("Sep") %>%
  recur_on_wday("Monday", nth = -1)

sch_seq("2000-01-01", "2005-01-01", on_last_monday_in_sept)
```

## Schedules

Recurrence rules are powerful tools on their own, but they aren't enough to solve every task. Say you want to construct a rule that includes both Christmas and Labor Day as events. It would be impossible to construct this kind of recurrence set using a single rule, but if you could join _multiple_ rules together, one for Christmas and one for Labor Day, then it would be possible.

A _schedule_ is a collection of recurrence rules. The recurrence set of a schedule is the union of the recurrence sets of every underlying rule. It can also contain specialized dates that you want to forcibly include in the recurrence set, called `rdate`s, and ones that you want to forcibly exclude, called `exdate`s.

```{r}
on_christmas <- yearly(since = "2000-01-01") %>%
  recur_on_ymonth("Dec") %>%
  recur_on_mday(25)

sch_christmas_or_labor_day <- schedule() %>%
  sch_add_rrule(on_christmas) %>%
  sch_add_rrule(on_labor_day)

sch_seq("2000-01-01", "2002-01-01", sch_christmas_or_labor_day)

sch_christmas_or_labor_day_except_2000_labor_day <- sch_christmas_or_labor_day %>%
  sch_add_exdate("2000-09-04")

sch_seq("2000-01-01", "2002-01-01", sch_christmas_or_labor_day_except_2000_labor_day)
```

Schedules are powerful because they allow us to begin to build up more complicated recurrence sets from simpler pieces. All of the `hldy_*()` holiday functions are schedules built up from simpler rules. As an example of how complex this can get, take a quick look at `hldy_thanksgiving()`. In present time, Thanksgiving takes place on the 4th Thursday of November, but it wasn't always this simple. The date that Thanksgiving occurs on has changed twice, which you can read about [here](https://www.archives.gov/legislative/features/thanksgiving) if you are interested, creating the following disjoint set of rules:

- Pre 1939: Last Thursday in November

- After 1939 but before 1942: Next to last Thursday in November

- Post 1942: The 4th Thursday in November

Here is the implementation of `hldy_thanksgiving()`, which combines these pieces into a larger schedule.

```{r}
hldy_thanksgiving
```

